Community-Conditioned Language Models

### 2020-11-19 paper/paper.tex
deal with some comments and other minor changes

@@

Second, we want to find out how, and how much, CCLMs can be attuned to
the specificity of communities. We test whether 1. the linguistic
models are correlated with a model of communities based only on
co-occurence of users between communities and 2. if latent community
classifiers based on the CCLM can accurately identify that a given
message comes from a given community. Additionally, one may
intuitively assume that linguistically poor communities are easier to
recognize: the less linguistic diversity, the more specific the
language become and the easier it is to identify.

To sum up, we propose to test the following hypotheses:
\begin{hypotheses}
\item \label{hyp:varying-complexity} Different communities have different levels of linguistic
  complexity
\item \label{hyp:layer-effect} The layer at which the community embedding is taken into account
  in a CCLM influences its perplexity.
\item \label{hyp:LMCC-works} CCLM can be used to recognize communities based on
  latent bayesian classification.
% \item \label{hyp:LMCC-works} The layer at which the community embedding is taken into account
%   influences CCLM and its associated latent classifier.
  % This is more of a justification for studying the latent classifier than a real "user-facing" hypothesis.
\item \label{hyp:extra-linguistic-correlation} The CCLM representation of communities are correlated with
  co-ocurrence of users between communities.
% \item Community stability is correlated with language complexity
\item \label{hyp:rich-harder-to-identify} Linguistically poor communities are easier to identify than
  linguistically rich communities.
\end{hypotheses}

@@

We experiment with two kinds of model architecture: LSTM \citep{Hochreiter1997}\jp{Is it really a simple or a bi-lstm? Even If it's simple it may be worth specifying due to the prevalence of bilstm everywhere.}
and Transformer \citep{Vaswani2017}.  In either case, the model
is organised as a standard $n$-layer neural sequence encoder with
token embedding (we set $n=3$ for all models). As is standard, words are predicted
using a softmax projection layer over a pre-determined vocabulary.
These encoders form the core of the community-conditioned language models (CCLM).

@@

We also note that the conditioned models perform differently across different communities---%
for some communities, perplexity of the conditioned models is a lot lower,\bn{is this ok? took ``test'' out but made clear in the caption that it's the test set}%\jp{and training?}
whereas for others it is the same or even slightly higher (\cref{tab:model-ppls}).\bn{talk about this in terms of mutual information instead?}
As an informal observation, we note 
that across all the models we tested, 
communities where conditioning has the least effect
tend to be more organized around more ``general interest'' topics,
such as /r/relationships, where the subject matter is relevant to a broad
range of people. \bn{I really mean this as a quality of the topic, rather than the community itself, although the two are probably correlated}% \jp{Is there a way to quantify general interest topics? Community size in terms of number of users?}
Conditioning the model on community has the most benefit
for narrower special-interest subreddits such as those organized
around a certain videogame, sports team, or subculture.
This makes sense intuitively, since communities with more niche topics
would tend to have more specialized vocabulary and speech patterns.\bn{i think it's ok---i've seen this broader sense used in academic papers}%\jp{can you say speech for written language?}

@@

We find that the layer at which the community embedding is taken into
account in a CCLM influences its perplexity (\cref{hyp:layer-effect}),
albeit weakly. For LSTM models, the perplexity per word, averaged over
all communities, varies between 76.80 and 78.64. \bn{80.70 is the baseline model} 
The perplexity per word is slightly higher for $l_c=0$, and has very
little difference between the other models, with $l_c=1$ being the best.
For transformer models, it varies
between 89.38 and 122.56. Here perplexity per-word is clearly not a monotonous
function of $l_c$: the maximum is reached at $l_c=2$. We did not
expect non-mononicity, and leave studying this effect to future work.

The pattern of information gain by community is similar across architectures.
Communities that bennifit most from the conditioned model behave that way for 
both the LSTM and transformer.
However, there are some differences.
For example, /r/relationships has an information gain of \num{1.018} for
the best CCLM LSTM but only \num{0.998} for the best transformer model. \bn{todo: community-wise CCLM info gain}
Why was the LSTM able to take advantage of the community information for
this community, where the transformer wasn't?
It would be interesting to investigate these differences in future work,
since it could reveal differences in the kind of linguistic variation the different
model architectures capture.

>>>>

\jp{We want to identify a niche that we want to fill. A quick summary of the state of the art + what is missing.}

@@

Second, we want to find out how, and how much, CCLMs are attuned to
the specificity of communities. We test whether 1. the linguistic
models are correlated with a model of communities based only on
co-occurence of users between communities and 2. if latent community
classifiers based on the CCLM can accurately identify that a given
message comes from a given community. Additionally, one may
intuitively assume that linguistically poor communities are easier to
recognize: the less linguistic diversity, the more specific the
language become and the easier it is to identify.

To sum up we aim at testing the following hypotheses:
\begin{hypotheses}
\item \label{hyp:varying-complexity} Different communities have different levels of linguistic
  complexity
\item \label{hyp:layer-effect} The layer at which the community embedding is taken into account
  in a CCLM influences its perplexity.
\item \label{hyp:LMCC-works} CCLM can be used to recognize communities based on
  latent bayesian classification.
% \item \label{hyp:LMCC-works} The layer at which the community embedding is taken into account
%   influences CCLM and its associated latent classifier.
  % This is more of a justification for studying the latent classifier than a real "user-facing" hypothesis.
\item \label{hyp:extra-linguistic-correlation} The CCLM representation of communities are correlated with
  co-ocurrence of users between communities.
% \item Community stability is correlated with language complexity
\item \label{hyp:rich-harder-to-identify} Linguistically poor communities are easier to identify than
  linguistically rich communities.
\end{hypotheses}

@@

We experiment with two kinds of model architecture: LSTM
\citep{Hochreiter1997}\jp{Is it really a simple or a bi-lstm? Even If
  it's simple it may be worth specifying due to the prevalence of
  bilstm everywhere.}  and Transformer \citep{Vaswani2017}.  In either
case, the model is organised as a standard $n$-layer neural sequence
encoder (we set $n=3$ for all models). As is standard, word tokens are
first converted to vectors using a trainable embedding layer. At the
other end, word tokens are predicted using a softmax projection layer. The set of word tokens is a
pre-determined vocabulary.  These encoders form the core of the
community-conditioned language models (CCLM), and we call them
unconditioned models.

@@

We also note that the conditioned models perform differently across different communities---%
for some communities, perplexity of the conditioned models is a lot lower than the unconditioned models,
whereas for others it is the same or even slightly higher (\cref{tab:model-ppls}).\bn{talk about this in terms of mutual information instead?}\jp{What is that?}
As an informal observation, we note 
that across all the models we tested, 
communities where conditioning has the least effect
tend to be more organized around more general interest topics,
such as the \emph{relationships} subreddit, where the subject matter is relevant to a broad
range of people.
Conditioning the model on community has the most benefit
for narrower special-interest subreddits such as those organized
around a certain videogame, sports team, or subculture.
This makes sense intuitively, since communities with more niche topics
would tend to have more specialized vocabulary and other linguistic patterns.

@@

We find that the layer at which the community embedding is taken into
account in a CCLM influences its perplexity (\cref{hyp:layer-effect}),
albeit weakly. For LSTM models, the perplexity per word, averaged over
all communities, varies between 76.80 and 78.64. (with 80.70 for the unconditioned model)
The perplexity per word is slightly higher for $l_c=0$, and has very
little difference between the other models, with $l_c=1$ being the best.
For transformer models, it varies
between 89.38 and 122.56. Here perplexity per-word is clearly not a monotonous
function of $l_c$: the maximum is reached at $l_c=2$. We did not
expect non-mononicity, and leave studying this effect to future work.

The pattern of information gain by community is similar across architectures.
Communities that bennifit most from the conditioned model behave that way for 
both the LSTM and transformer.
However, there are some differences.
For example, the subreddit \emph{relationships} has an information gain of \num{1.018} for
the best CCLM LSTM but only \num{0.998} for the best transformer model. \bn{todo: community-wise CCLM info gain}
Why was the LSTM able to take advantage of the community information for
this community, where the transformer wasn't?
It would be interesting to investigate these differences in future work,
since it could reveal differences in the kind of linguistic variation the different
model architectures capture.
